State of the Practice of Lattice Boltzmann Method Software Development

### 2021-06-02 StateOfPractice/Peter-Notes/Report/Report.tex
1st edits to results section of LBM Report

This section presents the AHP results for qualities that were quantitatively measured. Qualitiative data from interviews with some developers is also found in this section. Working quality definitions are found in section \ref{softwarequalities}.

@@

All of the software packages that were tested have installation instructions. As noted previously, many of the 23 software packages removed from the original list of 46 software packages are missing documentation, including installation instructions.

All 23 packages can be installed on some Unix-like systems. Seven packages could be installed on Windows, and 5 on macOS. Operating system compatibility is found in the documentation of 19 software packages. For this state of the practice assessment, all but one of the software packages were tested on Ubuntu. The remaining package was tested on CentOS.

@@

All but two of the software packages have an automation procedure for at least part of the installation process. Most of these packages use make to automate the installation. A couple of packages use custom scripts for automation.

Errors encountered during the installation process were sometimes quickly fixed thanks to descriptive error messages. At other times, error messages were vague and difficult to troubleshoot. Fourteen software packages definitively broke during installation. A couple of packages did not provide a definitive message of the success or failure of installation. In these instances validating the installation required performing a tutorial or running a script, as described below, if they were available. The installation of most software packages could be recovered if a descriptive error message was displayed. Only three software packages that displayed a descriptive error message were not recoverable, and most of these instances were due to hardware and operating system incompatibility, such as the requirement of CUDA.

@@

Sixteen software packages require less than 10 dependencies to be installed. All but one software package require less than 20 dependencies. Some packages may automatically install additional dependencies in the background. Eighteen of the software packages do not explicitly indicate software dependency versions. Some software package installation issues, specifically those occurring when manual installation of dependencies is required, may be avoided if the dependency versions are specified. Fifteen software packages do not have detailed instructions for installing dependencies.

Sixteen software packages have less than ten manual installation steps. If commands for the installation of dependencies is streamlined then none of the software packages take more than 20 steps to install. The average number of steps is about 8, and the fewest is 3.

All but six of the software packages have a way to verify the installation. Twelve have some sort of tutorial examples that can be run by the user. Most of these generate output that can be compared to the documentation. Some packages do not have a way to verify the output of the tutorials, and the user is left to assume correctness if there is no error prompted during the tutorial. Some other ways of installation validation include validation scripts, automatic validation after the installation, and instructions to manually review the file system.

Instructions for uninstallation is easily found for only one of the software packages.

@@

Figure \ref{Fig_Installability} shows the installability ranking of the software packages using AHP. Software packages with a higher score tend to have one set of linear installation instructions that is written as if the person doing the installation has none of the dependencies installed. The instructions often have compatible operating system versions listed and include instructions for the installation of dependencies. The top ranked packages also often incorporate some sort of the automation in the installation process and have fewer manual installation steps. The number of dependencies a package has does not correlate with a higher score. Installation validation, often through test examples with expected output, is correlated with a higher score.

Many software packages would benefit from a rewrite or reorganization of installation instructions. A single location for installation instructions would improve their maintainability and correctness. Listing compatible operating system and dependency versions would decrease installation time and errors, as would adding instructions on installing dependencies. Errors that may be encountered during the installation process need to prompt detailed messages or logs. Once a software package is installed, either a transparent automatic validation needs to be performed or the user needs to be able to perform a manual validation using test examples and output. Finally, uninstallation instructions should be included in the documentation.

>>>>

This section presents the quantitative and qualitative findings of this study.  The qualitative data that is presented was gathered through interviews with some developers. AHP results for qualities that were quantitatively measured are also found in this section. Quality definitions are found in section \ref{softwarequalities}.

@@

All of the software packages that were tested have installation instructions. As noted previously, many of the twenty three software packages that were part of the original long list of forty six software packages but were later removed do not include documentation or installation instructions.

All twenty three packages on the short list can be installed on some Unix-like systems. Seven packages could be installed on Windows, and five on macOS. Operating system compatibility is found in the documentation of nineteen software packages. All but one of the software packages were tested on Ubuntu for this state of the practice assessment. The remaining package was tested on CentOS.

@@

All but two of the software packages have automated some of the installation process. Most of these packages use Make to automate the installation, and a few of them use custom scripts.

Errors encountered during the installation process were often quickly fixed thanks to descriptive error messages. Systems that provided vague error messages were more difficult to troubleshoot. Only three software packages that displayed a descriptive error message were not recoverable, and most of these instances were due to hardware and operating system incompatibility, such as the requirement of CUDA. Fourteen software packages definitively broke during installation. A couple of packages did not provide a definitive message of the success or failure of installation. In these instances, validating the installation required performing a tutorial or running a script, as described below, if these were available.

@@

Sixteen software packages require less than ten dependencies to be installed. All but one software package require less than twenty dependencies. Some packages may automatically install additional dependencies in the background. Eighteen of the software packages do not explicitly indicate software dependency versions. Some software package installation issues, specifically those occurring when manual installation of dependencies is required, may be avoided if versions of dependencies are specified. Fifteen software packages do not have detailed instructions for installing dependencies.

Sixteen software packages have less than ten manual installation steps. If dependencies are installed in one command then none of the software packages take more than twenty steps to install. The average number of steps is about eight, and the fewest is three.

All but six of the software packages have a way to verify the installation. Twelve have some sort of tutorial examples that can be run by the user. Most of these tutorials provide expected output. Some packages do not have a way to verify the output of the tutorials, and the user is left to assume correctness if no error message is encountered. Some other ways of installation validation include validation scripts, automatic validation after the installation, and instructions to manually review the file system.

Uninstallation instructions were found for only one of the software packages.

@@

Figure \ref{Fig_Installability} shows the installability ranking of the software packages using AHP. Software packages with a higher score tend to have one set of linear installation instructions that are written as if the person doing the installation has none of the dependencies installed. The instructions often list compatible operating system versions and include instructions for the installation of dependencies. The top ranked packages often incorporate some sort of automation of the installation process and have fewer manual installation steps. The number of dependencies a package has does not correlate with a higher score. The ability to validate the installation process, often through tutorials or test examples that include expected output, is correlated with a higher score.

Many software packages would benefit from a rewrite or reorganization of installation instructions. A single location for installation instructions would improve their maintainability and correctness. Listing compatible operating system and dependency versions would decrease installation time and errors, as would adding instructions on installing dependencies. Installation process errors should prompt the system to display detailed messages. Once a software package is installed, either an automatic validation needs to be performed or the user needs to be able to perform a manual validation using test examples that include expected output. Finally, uninstallation instructions should be included in the documentation.
